Cosmology with photometric weak lensing surveys: constraints with redshift tomography of convergence peaks and moments

### 2016-05-02 photoz_paper/paper.tex
incorporated latest edits

\title{Cosmology with weak lensing photometric surveys: constraints with redshift tomography of non--Gaussian statistics}

@@

\begin{abstract}
Weak gravitational lensing is becoming a mature technique to constrain cosmological parameters, and future surveys will be able to constrain the dark energy equation of state $w$. When analyzing galaxy surveys, redshift information has proven to be a valuable addition to angular shear correlations. We forecast parameter constraints on the triplet $(\Omega_m,w,\sigma_8)$ for an LSST like galaxy photometric survey, using tomography of the shear-shear power spectrum, convergence peak counts and higher convergence moments. We find that redshift tomography with the power spectrum shrinks the $1\sigma$ confidence interval area in $(\Omega_m,w)$ space by a factor of 8 with respect to the case of a single redshift bin. We also find that adding non-Gaussian information from the peak counts and higher moments of the convergence further shrinks the constrained area in $(\Omega_m,w)$ by a factor of respectively 3 and 4. When we add parameter priors from Planck to our analysis, we find that the improvements using tomography and non-Gaussian statistics are less important, spanning a factor of 3 at most using power spectrum tomography, but only improving by 30\% when adding the peak counts. 
We evaluate the effect of uncorrected photometric redshift systematics on parameter constraints. We find that different statistics lead to different bias directions in parameter space, suggesting the possibility of eliminating this bias via self--calibration. 
\end{abstract}

@@

The measured galaxy ellipticity $\pmb{\epsilon}$ is an estimate of the cosmic shear $\pmb{\gamma}$ due to large scale structure if the non--lensed galaxy shape is a circle. Because the galaxies have intrinsic non circular shapes, the measured galaxy ellipticity $\pmb{\epsilon}_{\rm m}$ is the sum of a cosmic shear term and the intrinsic ellipticity (galaxy shape noise) \citep{wlreview}
\begin{equation}
\pmb{\epsilon}_{\rm m} = \pmb{\gamma} + \pmb{\epsilon}_{\rm n}
\end{equation} 
%
where $\pmb{\epsilon}_{\rm n}$ is a random Gaussian variable with zero mean and redshift dependent variance $\sigma_n(z_s)=0.15+0.035z_s$. This is equivalent to saying that the cosmic shear inferred from ellipticity observations $\pmb{\gamma}_{\rm m}$ can be written as the sum of the true cosmic shear plus a noise term $\pmb{\gamma}_{\rm n}$ with the same statistical properties as $\pmb{\epsilon}_{\rm n}$. 
We add independent random realizations of the shape noise $\pmb{\gamma}_{\rm n}$ to each of the $N_r$ shear catalogs. Each shape noise realization is generated with a different random seed. The same random seeds are used to generate shape noise catalogs across simulations with different cosmological parameters $\{\bb{p}_i\}$.

@@

The study of photometric redshift errors is an active subject of research, and includes investigation of techniques such as spectroscopic calibration, catastrophic errors and cross-correlation techniques that we do not explore in this work (see for example \citep{HuTomo} for a more thorough discussion). We model the effect of photo-$z$ errors as a constant bias term $b_{\rm ph}(z_s)$ plus a random Gaussian component with variance $\sigma_{\rm ph}(z_s)$,
\begin{equation}
\label{meth:photoz-correction}
z_p(z_s) = z_s + b_{\rm ph}(z_s) + \sigma_{\rm ph}(z_s)\mathcal{N}(0,1),   
\end{equation}
%
where $\mathcal{N}(0,1)$ is the standard normal distribution. We bin the $N_g$ galaxies in our simulated catalogs into 5 redshift bins $\bar{z}_b$, $b=1...5$. Several models have been proposed in the literature for the photometric bias $b_{\rm ph}(z_s)$ (see for example \citep{Huterer2006}) and variance $\sigma_{\rm ph}(z_s)$ (see for example \citep{LSSTSciBook}). We chose the photo-$z$ bias and variance functions in equation (\ref{meth:photoz-correction}) to be the science requirements contained in the LSST Science Book \citep{LSSTSciBook}, namely $b(z_s)=0.003(1+z_s)$ and $\sigma(z_s)=0.02(1+z_s)$.

@@

\subsection{Dimensionality reduction}
We apply a Principal Component Analysis (see \citep{astroMLText} for example) to reduce the dimensionality of our summary statistics while preserving the cosmological information content. The model statistic $\bb{d}(\bb{p})$ can be regarded as a$P\times N_b$ matrix $d_{pi}$. Consider the whitened model matrix
\begin{equation}
\label{meth:whitening}
\Delta_{pi} = P\frac{d_{pi}}{\sum_{q=1}^Pd_{qi}} - 1.
\end{equation} 
%
We perform a Singular Value Decomposition (SVD) of $\Delta$ 
\begin{equation}
\bb\Delta = \bb{L}\bb{\Lambda} \bb{R}
\end{equation}
%
where $L$ is $P\times Q$, $\bb{\Lambda}={\rm diag}(\Lambda_1,...,\Lambda_Q)$ and $\bb{R}$ is $Q\times N_b$ and $Q={\min}(P,N_b)$. $R_{ni}$ is the $i$-th component of the $n$-th basis vector in statistics space. The singular value $\Lambda_n$ is the variance of the whitened summary statistic along the $n$-th basis vector. We assume that only summary statistic projections on the first $N_c$ basis vectors contain relevant cosmological information, where $N_c<N_b$ is a number that has to be determined from the simulations. Let $\bb{R}(N_c)$ be a matrix made of the first $N_c$ rows of $\bb{R}$ (we assume that the singular values $\Lambda_i$ are sorted). We define the PCA projection of a summary statistic $\bbh{d}$ on $N_c$ principal components as 
\begin{equation}
\label{meth:pcaprojection}
\h{d}_n(N_c) = \sum_{i=1}^{N_b}(\bb{R}(N_c))_{ni}\left(P\frac{\h{d}_i}{\sum_{p=1}^P d_{pi}}-1\right) 
\end{equation} 
%
Through the above procedure, we hope to capture the cosmological information contained in $\bbh{d}$ by projecting it on the $N_c<N_b$ principal components that vary the most with cosmology parameters. In the next section we outline the main results of this work.

@@

\subsection{Priors from CMB experiments}
In this paragraph we describe how we included prior knowledge of cosmological parameters from previous Cosmic Microwave Background (CMB) observations, such as Planck \citep{PlanckCosmo}. CMB experiments provide tight constraints on $(\Omega_m,\sigma_8)$, but they are not sensitive to dark energy parameters such as $w$. Nevertheless, prior knowledge of $\Omega_m$ and $\sigma_8$ could in principle help in breaking degeneracies between these parameters and $w$ in weak lensing observations. The CMB parameter prior probability function can be written as 
\begin{equation}
-2\log\mathcal{L}_{\rm CMB}(\bb{p}) = (\bb{p} - \bb{p}_0)^T\bb{F}_{\rm CMB}(\bb{p} - \bb{p}_0)
\end{equation}
%
where we assumed that the best fit parameters are the same $\bb{p}_0$ that appear in equation (\ref{meth:linapprox}). 
Parameter constraints from the Planck CMB experiment are made available to the public via the parameter Markov Chains (MCMC) published on the Planck Legacy Archive \footnote{The archive we used is located \url{http://pla.esac.esa.int/pla/}; we used the MCMC chains contained in the \ttt{base\_w/plikHM\_TT\_lowTEB} directory, labeled as \ttt{base\_w\_plikHM\_TT\_lowTEB\_[1-4].txt}}. We can use these MCMC data to estimate the parameter covariance matrix $\bb{\Sigma}_{\rm CMB}$ on the parameter multiplet $(\Omega_m,\Omega_bh^2,\theta_{\rm MC},\tau,w,n_s,\sigma_8)$, marginalized over the Planck nuisance parameters. We then compute the parameter prior Fisher matrix $\bb{F}_{\rm CMB} = \bb{\Sigma}_{\rm CMB}^{-1}$. Fixing the values of all parameters but $(\Omega_m,w,\sigma_8)$ and applying the prior to the weak lensing parameter likelihood (\ref{meth:paramlikelihood}) is equivalent to taking the $(\Omega_m,w,\sigma_8)$ slice of $\bb{F}_{\rm CMB}$, which we call $\bb{F}_{\rm CMB}^{(\Omega_m,w,\sigma_8)}$, and computing the parameters constraints subject to the CMB prior as

@@

Figure \ref{fig:photozbias} shows the effect of ignoring photo-$z$ errors on parameter constraints. To evaluate this effect we construct different simulated observations, with and without photo-$z$ errors, and compare the results of the parameter fit according to equation (\ref{meth:parestimate}). Using our simulation suite, we construct 20 simulated observations: the summary statistic in each observation is calculated by taking the mean of a random sample of $N_{\rm FOV}$ realizations of the summary statistic in the fiducial cosmology (randomly chosen among the ensemble of $N_r=16000$ that are available in the ensemble). The estimated covariance matrix $\bbh{C}$ is scaled by a factor $N_{\rm FOV}$ to take into account the construction process of the simulated observations. This procedure allows us to forecast the results for an LSST-like survey would obtain.

\begin{figure*}
\includegraphics[scale=0.3]{Figures/constraints_Om-w_lensing.eps}
\includegraphics[scale=0.3]{Figures/constraints_Om-w_lensing_cmb.eps}
\caption{$1\sigma$ confidence contours on the $(\Omega_m,w)$ parameter space, marginalized over $\sigma_8$, obtained using equation (\ref{meth:parcovestimator}). The covariance matrix $\bbh{C}$ and its inverse $\bbh{\Psi}$ have been computed from $16000$ summary statistics realizations, and have been scaled by a factor $N_{\rm FOV}=1600$ to mimic the constraining power of an LSST-sized survey. The thick line ellipses in the right panel refer to $(\Omega_m,w)$ obtained from the weak lensing statistics considered in this work, but subject to Planck priors as described in equation (\ref{meth:parcovestimator_cmb}). The thin solid lines in the left and right panels are the same.}
\label{fig:constraintsOm-w}
\end{figure*}

@@

           Power spectrum + peaks ($\bar{z}_5$) &            0.0040 &     0.0209 &            0.0044 &                            58 &                                         53 \\
                  Power spectrum + peaks (tomo) &            0.0021 &     0.0153 &            0.0026 &                            27 &                                         26 \\
         Power spectrum + moments ($\bar{z}_5$) &            0.0023 &     0.0190 &            0.0025 &                            32 &                                         26 \\
                Power spectrum + moments (tomo) &            0.0016 &     0.0150 &            0.0019 &                            18 &                                         14 \\
 Power spectrum + peaks + moments ($\bar{z}_5$) &            0.0020 &     0.0127 &            0.0024 &                            21 &                                         17 \\
        Power spectrum + peaks + moments (tomo) &            0.0015 &     0.0121 &            0.0018 &                            14 &                                         11 \\ \hline
\bottomrule
\end{tabular}
\caption{Constraints on the $(\Omega_m,w,\sigma_8)$ parameter triplet using different summary statistics and redshift information. Each column $(\Delta \Omega_m,\Delta w,\Delta \sigma_8)$ contains the 1$\sigma$ constraints on a particular parameter, marginalized over the other two. The last two columns contain the respectively the area of the $(\Omega_m,w)$ 68\% confidence level ellipse (marginalized over $\sigma_8$) and the volume of the 68\% confidence level ellipsoid in $(\Omega_m,w,\sigma_8)$ space, both calculated as the square root of the determinant of the relevant $\bbh{\Sigma}$ minors.}
\label{tbl:constraints}
\end{table*}

@@

           Power spectrum + peaks ($\bar{z}_5$) &            0.0040 &     0.0184 &            0.0043 &                            40 &                                         36 \\
                  Power spectrum + peaks (tomo) &            0.0021 &     0.0127 &            0.0025 &                            20 &                                         19 \\
         Power spectrum + moments ($\bar{z}_5$) &            0.0022 &     0.0145 &            0.0025 &                            22 &                                         17 \\
                Power spectrum + moments (tomo) &            0.0015 &     0.0120 &            0.0018 &                            14 &                                         11 \\
 Power spectrum + peaks + moments ($\bar{z}_5$) &            0.0019 &     0.0110 &            0.0023 &                            16 &                                         13 \\
        Power spectrum + peaks + moments (tomo) &            0.0015 &     0.0104 &            0.0018 &                            12 &                                          9 \\ \hline
\bottomrule
\end{tabular}
\caption{Constraints on the $(\Omega_m,w,\sigma_8)$ parameter triplet using different summary statistics and redshift information, including Fisher priors from Planck according to equation (\ref{meth:parcovestimator_cmb}). Each column $(\Delta \Omega_m,\Delta w,\Delta \sigma_8)$ contains the 1$\sigma$ constraints on a particular parameter, marginalized over the other two. The last two columns contain the respectively the area of the $(\Omega_m,w)$ 68\% confidence level ellipse (marginalized over $\sigma_8$) and the volume of the 68\% confidence level ellipsoid in $(\Omega_m,w,\sigma_8)$ space, both calculated as the square root of the determinant of the relevant $\bbh{\Sigma}$ minors.}
\label{tbl:constraints-cmb}
\end{table*}

@@

Figure \ref{fig:nopca} also clearly shows that, when considering a single redshift bin and a single summary statistic, most of the information on $w$ is contained in the highest redshift galaxies. PCA does not seem to capture the whole information in the lowest redshift bin, even when enough components (enough for the plateau in Figure \ref{fig:pcacomponents} to be reached) are included. This can be attributed to the fact that PCA is not scale-invariant \citep{astroMLText}, because there is freedom in choosing the whitening scale in equation (\ref{meth:whitening}). Our choice of the whitening scale seems to affect significantly the first redshift bin, with the effect being mitigated on the highest redshift bins. 
The top left panel of Figure \ref{fig:nopca} also shows reasonable agreement between the results we obtain with our simulations and the ones we calculate with the analytical code NICAEA.

@@

Table \ref{tbl:constraints-cmb} shows that parameter priors on Planck yield a factor of 6 improvement on the $(\Omega_m,w)$ and $(\Omega_m,w,\sigma_8)$ 68\% confidence intervals, even when a single redshift bin is considered. When the Planck priors in equation (\ref{meth:parcovestimator_cmb}) are included, the improvements in constraining power when adding redshift tomography or combining different statistics are more moderate, with power spectrum tomography yielding a factor of 3 improvement, peaks tomography yielding no improvement and moments tomography yielding a factor 1.5 improvement.

@@

	\item When considering a single redshift bin, most of the cosmological information on $w$ is contained in high redshift galaxies. Constraints can be improved with redshift tomography or combining different non-Gaussian statistics with the power spectrum. The improvement originating from the combination of different statistics is attributed to the complementary information that non-Gaussian statistics carry, as a similar improvement cannot be obtained simply increasing the number of PCA components in a single statistic.

	\item Redshift tomography on the power spectrum shrinks the $(\Omega_m,w)$ 68\% confidence ellipse by a factor of 8; combining the peak counts with the power spectrum in the highest redshift bin leads to a factor of 10 better constraints, while adding the moments instead reduces the size of the $(\Omega_m,w)$ ellipse by a factor of 20. When redshift tomography is added on top of these statistics combinations, an additional factor of 2 improvement is observed. Constraint improvements adding redshift tomography and combinations of different statistics are less important when priors from CMB experiments are included in the analysis.

@@

This work emphasizes the advantage of deep galaxy surveys such as LSST, which have access to shape information of high redshift galaxies and provide valuable cosmological information on the dark energy equation of state. We also stress the fact that redshift tomography can in some cases provide more stringent constraints on parameters but, for this technique to be viable, accurate knowledge of galaxy redshifts is necessary. Future work needs to address the requirements for photometric measurements accuracy when using non-Gaussian statistics, as well as the self calibration techniques that can be used when different summary statistics are available in addition to the power spectrum.

@@

We thank Colin Hill and Licia Verde for useful discussions. We also thank Martin Kilbinger, Lloyd Knox and Hu Zhan for comments on an earlier version of this manuscript.
The simulations in this work were performed at the NSF XSEDE facility, supported by grant number
ACI-1053575, and at the New York Center for Computational Sciences, a
cooperative effort between Brookhaven National Laboratory and Stony
Brook University, supported in part by the State of New York. This
work was supported in part by the U.S. Department of Energy under
Contract Nos. DE-AC02-98CH10886 and DE-SC0012704, and by the NSF Grant
No. AST-1210877 (to Z.H.) and by the Research Opportunities and
Approaches to Data Science (ROADS) program at the Institute for Data
Sciences and Engineering at Columbia University (to Z.H.).

\textbf{To Morgan's attention: the bulk of the simulations were performed at NERSC, I do not know under what terms the allocation was granted though. Can you fill the appropriate details?}

>>>>

\title{Cosmology with photometric weak lensing surveys: constraints with redshift tomography of convergence peaks and moments}

@@

\begin{abstract}
Weak gravitational lensing is becoming a mature technique for constraining cosmological parameters, and future surveys will be able to constrain the dark energy equation of state $w$. When analyzing galaxy surveys, redshift information has proven to be a valuable addition to angular shear correlations. We forecast parameter constraints on the triplet $(\Omega_m,w,\sigma_8)$ for an LSST-like galaxy photometric survey, using tomography of the shear-shear power spectrum, convergence peak counts and higher convergence moments. We find that redshift tomography with the power spectrum shrinks the $1\sigma$ confidence interval area in $(\Omega_m,w)$ space by a factor of 8 with respect to the case of the single highest redshift bin. We also find that adding non-Gaussian information from the peak counts and higher moments of the convergence further shrinks the constrained area in $(\Omega_m,w)$ by a factor of respectively 3 and 4. When we add parameter priors from Planck to our analysis, tomography improves power spectrum constraints by a factor of 3. Adding moments improves by an additional factor of 2, and  adding both moments and peaks improves by almost a factor of 3 over power spectrum tomography alone. We evaluate the effect of uncorrected photometric redshift systematics on parameter constraints. We find that different statistics lead to different bias directions in parameter space, suggesting the possibility of eliminating this bias via self-calibration. 
\end{abstract}

@@

The measured galaxy ellipticity $\pmb{\epsilon}$ is an estimate of the cosmic shear $\pmb{\gamma}$ due to large scale structure if the non--lensed galaxy shape is a circle. Because the galaxies have intrinsic noncircular shapes, the measured galaxy ellipticity $\pmb{\epsilon}_{\rm m}$ is the sum of a cosmic shear term and the intrinsic ellipticity (galaxy shape noise) \citep{wlreview}
\begin{equation}
\pmb{\epsilon}_{\rm m} = \pmb{\gamma} + \pmb{\epsilon}_{\rm n}
\end{equation} 
%
where $\pmb{\epsilon}_{\rm n}$ is a random Gaussian variable with zero mean and redshift dependent variance $\sigma_n(z_s)=0.15+0.035z_s$. This is equivalent to saying that the cosmic shear inferred from ellipticity observations $\pmb{\gamma}_{\rm m}$ can be written as the sum of the true cosmic shear plus a noise term $\pmb{\gamma}_{\rm n}$ with the same statistical properties as $\pmb{\epsilon}_{\rm n}$. 
We add independent random realizations of the shape noise $\pmb{\gamma}_{\rm n}$ to each of the $N_r$ shear catalogs. Each shape noise realization is generated with a different random seed. The same random seeds are used to generate shape noise catalogs across simulations with different cosmological parameters $\{\bb{p}_i\}$.

@@

The study of photometric redshift errors is an active subject of research, and includes investigation of techniques such as spectroscopic calibration, catastrophic errors and cross-correlation techniques that we do not explore in this work (see for example \citep{HuTomo,LSSTSciBook} for a more thorough discussion). We model the effect of photo-$z$ errors as a constant bias term $b_{\rm ph}(z_s)$ plus a random Gaussian component with variance $\sigma_{\rm ph}(z_s)$,
\begin{equation}
\label{meth:photoz-correction}
z_p(z_s) = z_s + b_{\rm ph}(z_s) + \sigma_{\rm ph}(z_s)\mathcal{N}(0,1),   
\end{equation}
%
where $\mathcal{N}(0,1)$ is the standard normal distribution. We bin the $N_g$ galaxies in our simulated catalogs into 5 redshift bins $\bar{z}_b$, $b=1...5$. Several models have been proposed in the literature for the photometric bias $b_{\rm ph}(z_s)$ (see for example \citep{Huterer2006}) and variance $\sigma_{\rm ph}(z_s)$ (see for example \citep{LSSTSciBook}). We chose the photo-$z$ bias and variance functions in equation (\ref{meth:photoz-correction}) to be the science requirements contained in the LSST Science Book \citep{LSSTSciBook}, namely $b(z_s)=0.003(1+z_s)$ and $\sigma(z_s)=0.02(1+z_s)$.

@@

\subsection{Dimensionality reduction}
We apply a Principal Component Analysis (see \citep{astroMLText} for example) to reduce the dimensionality of our summary statistics while preserving the cosmological information content. The model statistic $\bb{d}(\bb{p})$ can be regarded as a $P\times N_b$ matrix $d_{pi}$. Consider the whitened model matrix
\begin{equation}
\label{meth:whitening}
\Delta_{pi} = P\frac{d_{pi}}{\sum_{q=1}^Pd_{qi}} - 1.
\end{equation} 
%
We perform a Singular Value Decomposition (SVD) of $\Delta$ 
\begin{equation}
\bb\Delta = \bb{L}\bb{\Lambda} \bb{R}
\end{equation}
%
where $L$ is $P\times Q$, $\bb{\Lambda}={\rm diag}(\Lambda_1,...,\Lambda_Q)$ and $\bb{R}$ is $Q\times N_b$ and $Q={\min}(P,N_b)$. $R_{ni}$ is the $i$-th component of the $n$-th basis vector in statistics space. The singular value $\Lambda_n$ is the variance of the whitened summary statistic along the $n$-th basis vector. We assume that only summary statistic projections on the first $N_c$ basis vectors contain relevant cosmological information, where $N_c<N_b$ is a number that has to be determined from the simulations. Let $\bb{R}(N_c)$ be a matrix made of the first $N_c$ rows of $\bb{R}$ (we assume that the singular values $\Lambda_i$ are sorted). We define the PCA projection of a summary statistic $\bbh{d}$ on $N_c$ principal components as 
\begin{equation}
\label{meth:pcaprojection}
\h{d}_n(N_c) = \sum_{i=1}^{N_b}(\bb{R}(N_c))_{ni}\left(P\frac{\h{d}_i}{\sum_{p=1}^P d_{pi}}-1\right) 
\end{equation} 
%
Through the above procedure, we hope to capture the cosmological information contained in $\bbh{d}$ by projecting it on the $N_c<N_b$ principal components that vary the most with cosmology parameters. In the next section we outline the main results of this work.

@@

\subsection{Priors from CMB experiments}
In this section we describe how we included prior knowledge of cosmological parameters from previous Cosmic Microwave Background (CMB) observations, such as Planck \citep{PlanckCosmo}. CMB experiments provide tight constraints on $(\Omega_m,\sigma_8)$, but they are not sensitive to dark energy parameters such as $w$. Nevertheless, prior knowledge of $\Omega_m$ and $\sigma_8$ could in principle help in breaking degeneracies between these parameters and $w$ in weak lensing observations. The CMB parameter prior probability function can be written as 
\begin{equation}
-2\log\mathcal{L}_{\rm CMB}(\bb{p}) = (\bb{p} - \bb{p}_0)^T\bb{F}_{\rm CMB}(\bb{p} - \bb{p}_0)
\end{equation}
%
where we assumed that the best fit parameters are the same $\bb{p}_0$ that appear in equation (\ref{meth:linapprox}). 
Parameter constraints from the Planck CMB experiment are made available to the public via the parameter Markov Chains (MCMC) published on the Planck Legacy Archive \footnote{The archive we used is located \url{http://pla.esac.esa.int/pla/}; we used the MCMC chains contained in the \ttt{base\_w/plikHM\_TT\_lowTEB} directory, labeled as \ttt{base\_w\_plikHM\_TT\_lowTEB\_[1-4].txt}}. We can use these MCMC data to estimate the parameter covariance matrix $\bb{\Sigma}_{\rm CMB}$ on the parameter multiplet $(\Omega_m,\Omega_bh^2,\theta_{\rm MC},\tau,w,n_s,\sigma_8)$, marginalized over the Planck nuisance parameters. We then compute the parameter prior Fisher matrix $\bb{F}_{\rm CMB} = \bb{\Sigma}_{\rm CMB}^{-1}$. Fixing the values of all parameters but $(\Omega_m,w,\sigma_8)$ and applying the prior to the weak lensing parameter likelihood (\ref{meth:paramlikelihood}) is equivalent to taking the $(\Omega_m,w,\sigma_8)$ slice of $\bb{F}_{\rm CMB}$, which we call $\bb{F}_{\rm CMB}^{(\Omega_m,w,\sigma_8)}$, and computing the parameters constraints subject to the CMB prior as

@@

Figure \ref{fig:photozbias} shows the effect of ignoring photo-$z$ errors on parameter constraints. To evaluate this effect we construct different simulated observations, with and without photo-$z$ errors, and compare the results of the parameter fit according to equation (\ref{meth:parestimate}). Using our simulation suite, we construct 20 simulated observations: the summary statistic in each observation is calculated by taking the mean of a random sample of $N_{\rm FOV}=1600$ realizations of the summary statistic in the fiducial cosmology (randomly chosen among the ensemble of $N_r=16000$ that are available in the ensemble). The estimated covariance matrix $\bbh{C}$ is scaled by a factor $N_{\rm FOV}$ to take into account the construction process of the simulated observations. This procedure allows us to forecast the results an LSST-like survey would obtain.

\begin{figure*}
\includegraphics[scale=0.3]{Figures/constraints_Om-w_lensing.eps}
\includegraphics[scale=0.3]{Figures/constraints_Om-w_lensing_cmb.eps}
\caption{$1\sigma$ tomographic constraints on the $(\Omega_m,w)$ parameter space, marginalized over $\sigma_8$, obtained using equation (\ref{meth:parcovestimator}). The covariance matrix $\bbh{C}$ and its inverse $\bbh{\Psi}$ have been computed from $16000$ summary statistics realizations, and have been scaled by a factor $N_{\rm FOV}=1600$ to mimic the constraining power of an LSST-sized survey. The thick line ellipses in the right panel refer to $(\Omega_m,w)$ obtained from the weak lensing statistics considered in this work, but subject to Planck priors as described in equation (\ref{meth:parcovestimator_cmb}). The thin solid lines in the left and right panels are the same.}
\label{fig:constraintsOm-w}
\end{figure*}

@@

           Power spectrum + peaks ($\bar{z}_5$) &            0.0040 &     0.0209 &            0.0044 &                            58 &                                         53 \\
                  Power spectrum + peaks (tomo) &            0.0021 &     0.0153 &            0.0026 &                            27 &                                         26 \\
         Power spectrum + moments ($\bar{z}_5$) &            0.0023 &     0.0190 &            0.0025 &                            32 &                                         26 \\
                Power spectrum + moments (tomo) &            0.0016 &     0.0150 &            0.0019 &                            18 &                                         14 \\
 Power spectrum + peaks + moments ($\bar{z}_5$) &            0.0020 &     0.0127 &            0.0024 &                            21 &                                         17 \\
        Power spectrum + peaks + moments (tomo) &            0.0015 &     0.0121 &            0.0018 &                            14 &                                         11 \\ \hline
\bottomrule
\end{tabular}
\caption{Constraints on the $(\Omega_m,w,\sigma_8)$ parameter triplet using different summary statistics and redshift information. Each column $(\Delta \Omega_m,\Delta w,\Delta \sigma_8)$ contains the 1$\sigma$ constraints on a particular parameter, marginalized over the other two. The last two columns contain respectively the area of the $(\Omega_m,w)$ 68\% confidence level ellipse (marginalized over $\sigma_8$) and the volume of the 68\% confidence level ellipsoid in $(\Omega_m,w,\sigma_8)$ space, both calculated as the square root of the determinant of the relevant $\bbh{\Sigma}$ minors.}
\label{tbl:constraints}
\end{table*}

@@

           Power spectrum + peaks ($\bar{z}_5$) &            0.0040 &     0.0184 &            0.0043 &                            40 &                                         36 \\
                  Power spectrum + peaks (tomo) &            0.0021 &     0.0127 &            0.0025 &                            20 &                                         19 \\
         Power spectrum + moments ($\bar{z}_5$) &            0.0022 &     0.0145 &            0.0025 &                            22 &                                         17 \\
                Power spectrum + moments (tomo) &            0.0015 &     0.0120 &            0.0018 &                            14 &                                         11 \\
 Power spectrum + peaks + moments ($\bar{z}_5$) &            0.0019 &     0.0110 &            0.0023 &                            16 &                                         13 \\
        Power spectrum + peaks + moments (tomo) &            0.0015 &     0.0104 &            0.0018 &                            12 &                                          9 \\ \hline
\bottomrule
\end{tabular}
\caption{Constraints on the $(\Omega_m,w,\sigma_8)$ parameter triplet using different summary statistics and redshift information, including Fisher priors from Planck according to equation (\ref{meth:parcovestimator_cmb}). Each column $(\Delta \Omega_m,\Delta w,\Delta \sigma_8)$ contains the 1$\sigma$ constraints on a particular parameter, marginalized over the other two. The last two columns contain respectively the area of the $(\Omega_m,w)$ 68\% confidence level ellipse (marginalized over $\sigma_8$) and the volume of the 68\% confidence level ellipsoid in $(\Omega_m,w,\sigma_8)$ space, both calculated as the square root of the determinant of the relevant $\bbh{\Sigma}$ minors.}
\label{tbl:constraints-cmb}
\end{table*}

@@

Figure \ref{fig:nopca} also clearly shows that, when considering a single redshift bin and a single summary statistic, most of the information on $w$ is contained in the highest redshift galaxies. PCA does not seem to capture all the information in the lowest redshift bin, even when enough components are included to reach the plateau in Figure \ref{fig:pcacomponents}. This can be attributed to the fact that PCA is not scale-invariant \citep{astroMLText}, because there is freedom in choosing the whitening scale in equation (\ref{meth:whitening}). Our choice of the whitening scale seems to affect significantly the first redshift bin, with the effect being mitigated for the highest redshift bins. 
The top left panel of Figure \ref{fig:nopca} also shows reasonable agreement between the results we obtain with our simulations and the ones we calculate with the analytical code NICAEA.

@@

Table \ref{tbl:constraints-cmb} shows that parameter priors from Planck yield a factor of 6 improvement on the $(\Omega_m,w)$ and $(\Omega_m,w,\sigma_8)$ 68\% confidence intervals, even when a single redshift bin is considered. When the Planck priors in equation (\ref{meth:parcovestimator_cmb}) are included, the improvements in constraints when adding redshift tomography or combining different statistics are more moderate. Tomography improves power spectrum constraints by a factor of 3. Adding moments improves by an additional factor of 2, and  adding both moments and peaks improves by almost a factor of 3 over power spectrum tomography alone.

@@

	\item When considering a single redshift bin, most of the cosmological information on $w$ is contained in high redshift galaxies. Constraints can be improved with redshift tomography or combining different non-Gaussian statistics with the power spectrum. The improvement originating from the combination of different statistics is attributed to the complementary information that non-Gaussian statistics carry, as a similar improvement cannot be obtained from a single statistic.

	\item Redshift tomography on the power spectrum shrinks the $(\Omega_m,w)$ 68\% confidence ellipse by a factor of 8; combining the peak counts with the power spectrum in the highest redshift bin leads to a factor of 10 better constraints, while adding the moments instead reduces the size of the $(\Omega_m,w)$ ellipse by a factor of 20. When redshift tomography is added on top of these statistics combinations, an additional factor of 2 improvement is observed. Constraint improvements adding redshift tomography and combinations of different statistics are less dramatic when priors from CMB experiments are included in the analysis.

@@

This work explores the advantage of deep galaxy surveys such as LSST, which have access to shape and redshift information of high $z$ galaxies and provide valuable cosmological information on the dark energy equation of state. We also stress the fact that redshift tomography can in some cases provide more stringent constraints on parameters but, for this technique to be viable, accurate knowledge of galaxy redshifts is necessary. Future work needs to address the requirements for photometric measurements accuracy when using non-Gaussian statistics, as well as the self calibration techniques that can be used when different summary statistics are available in addition to the power spectrum.

@@

Most of the calculations were performed at National Energy Research Scientific Computing Center (NERSC). We thank the LSST Dark Energy Science Collaboration (DESC) for the allocation of time, and for many useful discussions. We thank Hu Zhan, Salman Habib, Jeffrey Newman, Deborah Bard, and Peter Nugent for technical help. We thank Colin Hill and Licia Verde for useful discussions. We also thank Martin Kilbinger and Lloyd Knox for comments on an earlier version of this manuscript.
Part of the simulations in this work were also performed at the NSF XSEDE facility, supported by grant number
ACI-1053575, and at the New York Center for Computational Sciences, a
cooperative effort between Brookhaven National Laboratory and Stony
Brook University, supported in part by the State of New York. This
work was supported in part by the U.S. Department of Energy under
Contract Nos. DE-AC02-98CH10886 and DE-SC0012704, and by the NSF Grant
No. AST-1210877 (to Z.H.) and by the Research Opportunities and
Approaches to Data Science (ROADS) program at the Institute for Data
Sciences and Engineering at Columbia University (to Z.H.).
